### 2021-03-31 ch01_introduction/introduction.tex
Standardized the usage of 'emph' and 'enquote'

To successfully mitigate the massive impacts of climate change, a first
important step is the understanding of climate change and its accurate
quantification. Extremely valuable tools in this context are climate models,
which allow us to simulate the behavior of the climate system under arbitrary
conditions without having to perform (ethnically questionable) experiments in
the real world. All around the world climate research institutes provide a
variety of different climate models. Many of them participate in the \ac{CMIP},
which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
\enquote{better understand past, present and future climate changes arising
  from natural, unforced variability or in response to changes in radiative
  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
provide crucial input for the international climate assessments given by the
\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
  as \acs{CMIP}6} support the assessment of the upcoming \ac{IPCC} Sixth
\acl{AR} (\acs{AR}6), and their predecessor models from \acs{CMIP}5
\autocite{Taylor2012} have been assessed as part of the Fifth \acl{AR}
(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
allow the simulation of biological and chemical processes in addition to the
dynamics of the physical components of the Earth system, are also known as
\acp{ESM} and provide the most sophisticated simulations of the Earth's
climate. In this thesis, the terms \enquote{climate model} and
\enquote{\acl{ESM}} are used interchangeably since most modern models
participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.

Simulations that extrapolate state of the climate system into the future are
called \emph{projections} of the future climate. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with large ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. A state-of-the-art technique for this is the \emph{emergent
  constraints} method, which uses a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is the weighting of climate
models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

>>>>

To successfully mitigate the massive impacts of climate change, a first
important step is the understanding of climate change and its accurate
quantification. Extremely valuable tools in this context are climate models,
which allow us to simulate the behavior of the climate system under arbitrary
conditions without having to perform (ethnically questionable) experiments in
the real world. All around the world climate research institutes provide a
variety of different climate models. Many of them participate in the \ac{CMIP},
which was initiated in 1995 by the \ac{WGCM} of the \ac{WCRP} to
\enquote{better understand past, present and future climate changes arising
  from natural, unforced variability or in response to changes in radiative
  forcing in a multi-model context} \autocite{WCRP2020}. The \ac{CMIP} models
provide crucial input for the international climate assessments given by the
\acp{AR} of the \ac{IPCC}. For example, the latest generation of climate models
from the most recent (sixth) phase of \ac{CMIP} \commentcite{Eyring2016}{known
  as \acs{CMIP}6} support the assessment of the upcoming \ac{IPCC} Sixth
\acl{AR} (\acs{AR}6), and their predecessor models from \acs{CMIP}5
\autocite{Taylor2012} have been assessed as part of the Fifth \acl{AR}
(\acs{AR}5) in 2013 \autocite{Flato2013}. Modern-day climate models, which
allow the simulation of biological and chemical processes in addition to the
dynamics of the physical components of the Earth system, are also known as
\emph{\acp{ESM}} and provide the most sophisticated simulations of the Earth's
climate. In this thesis, the terms \enquote{climate model} and
\enquote{\acl{ESM}} are used interchangeably since most modern models
participating in \ac{CMIP} are \acp{ESM} or at least have \ac{ESM} versions.

Simulations that extrapolate the state of the climate system into the future
are called \emph{climate model projections}. These include idealized
simulations with only a prescribed change in the atmospheric \ac{CO2}
concentration (\eg{} an instantaneous doubling of \ac{CO2} or a \ac{CO2}
increase of $1 \unit{\%}$ per year) as well as more realistic projections that
consider different future scenarios (\eg{} a fossil fuel--based future or a
scenario that is based on a sustainable development). In many variables that
are relevant for climate change, multi-model projections from \ac{CMIP} show a
large inter-model range \autocite{Collins2013, Flato2013}. A crucial and
policy-relevant example for this is the climate sensitivity, which refers to
the change in the \ac{GSAT} that results from a change in the radiative
forcing. Common metrics for this are the \emph{\ac{ECS}}, which describes the
equilibrium response of the climate system after a doubling of the atmospheric
\ac{CO2} concentration, and the \emph{\ac{TCR}}, which describes the transient
response of the system to a \ac{CO2} doubling. In \acs{AR}5, both quantities
have been assessed with large ranges of $\rangeunit{1.5}{4.5}{K}$ and
$\rangeunit{1.0}{2.5}{K}$ for \ac{ECS} and \ac{TCR}, respectively
\autocite{Stocker2013}. The corresponding inter-model ranges from the
\acs{CMIP}5 models show similar results \autocite{Flato2013}. For this reason,
a careful statistical evaluation and further refinement of the output of
multi-model climate projections is necessary in order to reduce associated
uncertainties. A state-of-the-art technique for this is the \emph{emergent
  constraints} method, which uses a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations
\autocite{Allen2002}. An alternative approach is the weighting of climate
models based on their performance (\ie{} the distance of one model to
observational products) and interdependence (\ie{} the distance of one model to
other climate models) \autocite{Knutti2017a}. These techniques form the
baseline for the new analyses and results presented in this thesis, which
partly utilize methods from a new emerging research field in climate sciences:
\ac{AI} and \ac{ML}.

### 2021-03-31 ch02_scientific_background/scientific_background.tex
Standardized the usage of 'emph' and 'enquote'

A possible way out is given by numerical climate models, which offer the
possibility to simulate the Earth's climate on a computer. To efficiently
replicate the vastly complex Earth system with finite computational resources,
climate models divide the Earth into a set of \emph{grid cells}. The typical
size of an atmospheric grid cell in modern-day global climate models is about
$100 \unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
\unit{km}$ along the vertical dimension (pressure level or height). In
addition, also the temporal evolution of the Earth system is discretized using
time steps that are typically about $30 \unit{min}$ long in modern-day global
climate models. For each grid cell and time step, usually a single value per
model variable is given. Examples for such variables provided by a climate
model are the atmospheric prognostic variables velocity (horizontal and
vertical), temperature, specific humidity, pressure, and density. These
prognostic variables are related to each other via the \emph{primitive
  equations}, a set of partial differential equations that can be derived from
the conservation of momentum, mass, energy, and moisture \autocite{Holton2004}.
To progress further in time, these primitive equations are solved numerically
by the dynamical core of the climate models, which eventually describes the
Earth's large-scale atmospheric motions. Similar to this, many other processes
of the Earth system are simulated by the climate models using other fundamental
physical laws and principles. As opposed to these calculated variables that
form the output of climate models, the corresponding input is mainly given by
the radiative forcing.

@@

Simulations from climate model ensembles of \ac{CMIP} allow us to assess future
climate change in a consistent and transparent way. Especially the
\ac{ScenarioMIP} experiments can give valuable insights into possible
developments of the Earth system by providing \emph{projections} of the future
climate. In contrast to climate predictions, climate projections run over
multiple decades and depend upon the future scenario considered, which are
based on assumptions that may or may not turn out to be correct. On the
contrary, climate predictions are attempts to predict the actual evolution of
the climate on much shorter time scales from seasons to years. Similar to any
other scientific experiment, climate model projections suffer from associated
uncertainties. There are three major sources of uncertainties in climate model
projections we can distinguish: natural variability, climate response
uncertainty, and emission uncertainty \autocite{Hawkins2009, Hawkins2010}.
\Cref{fig:02:sources_of_uncertainty} shows these three sources for the
projected global mean surface temperature anomaly over the \nth{21} century.

@@

\emph{Natural variability} is connected to the chaotic nature of the Earth
system that arises from complex interactions between the ocean, atmosphere,
land, biosphere, and cryosphere \autocite{Cubasch2013}. It constitutes a
fundamental limit of how precisely we can project the future climate since it
is inherent in the Earth system and cannot be eliminated by more knowledge and
more advanced climate models. Natural variability is more relevant on regional
and local scales than on continental or global scales. Further contributions to
natural variability on longer time scales come from phenomena like the
\ac{ENSO} or the \ac{NAO} and from other events like volcanic eruptions and
variations in the solar activity. Natural variability can be seen as the
\enquote{noise} in the climate record as opposed to the anthropogenic
\enquote{signal} \autocite{Cubasch2013}.

@@

A commonly used distance metric to measure model performance $D_m$ and model
interdependence $S_{mn}$ is the \ac{RMSE}, but others are possible
\autocite{Knutti2017a}. The metrics are evaluated on a set of past or
present-day diagnostics and variables, whose choice is crucial for the
weighting scheme. A helpful strategy for this is to focus on addressing the
question \enquote{which climate model is adequate for predicting the target
  variable $y$?} instead of trying to answer the question \enquote{which
  climate model is the best?} \autocite{Parker2009}. Thus, diagnostics and
variables are chosen that are relevant for the projection of the target
variable \autocite{Knutti2017a}. In practice, this choice is either based on
expert judgment about relevant processes, on emergent relationships (see
\cref{subsec:02:emergent_constraints}) or on multivariate regression models
\commentcite{Karpechko2013, Sanderson2015a}{see \cref{subsec:02:mder}}. It
might also be beneficial to use different diagnostics for the calculations of
the performance and interdependence metrics \autocite{Merrifield2020} and/or to
remove selected diagnostics based on their mutual correlation
\autocite{Lorenz2018}.

The constants $\sigma_D$ and $\sigma_S$ determine how strongly the climate
models' performance and interdependence are weighted \autocite{Knutti2017a}.
Small values of the performance parameter $\sigma_D$ lead to an aggressive
weighting with only a few climate models receiving a majority of the weight,
while large values of $\sigma_D$ result in an equal weighting. For the
interdependence parameter $\sigma_S$, this is slightly different: Here, small
(all climate models are independent) and large (all climate models are
dependent) values lead to an almost equal weighting. Thus, an optimal choice
for $\sigma_D$ and $\sigma_S$ is crucial. A useful tool to estimate these
optimal parameters is the \emph{leave-one-model-out \ac{CV}} approach, which is
also known as \emph{pseudo-reality}, \emph{model-as-truth} or \emph{perfect
  model} setup \autocite{Elia2002, Karpechko2013}. For this, a single climate
model is removed from the multi-model ensemble and treated as observation
(\emph{pseudo-observation}). Then, a weighted \ac{MMM} with weights computed
from the updated model ensemble is calculated, which gives a prediction for the
\enquote{true} climate model. This allows a simple quantitative assessment of
the weighting scheme by calculating the \ac{RMSE} between the prediction and
the known ground truth of the pseudo-observation, which is also known as
\emph{\ac{RMSEP}}. The whole process is repeated for every climate model of the
ensemble to get a statistical distribution of \acp{RMSEP}. Finally, different
\ac{RMSEP} distributions calculated from different parameters $\sigma_D$ and
$\sigma_S$ can be assessed using specific criteria to find optimal values for
$\sigma_D$ and $\sigma_S$ \autocite{Knutti2017a}. Furthermore, the
leave-one-model-out \ac{CV} approach can be used to evaluate different climate
model weightings schemes (including the unweighted \ac{MMM}) and compare them
against each other.

@@

The concept of \ac{MDER} is mathematically similar to the concept of emergent
constraints. In a first step, an inter-model relationship between the target
variable and the process-relevant diagnostics is used to fit a multivariate
linear regression model. Let $\bm{y} = \left( y_1, y_2, \ldots, y_M \right)^T
\in \mathbb{R}^M$ be the vector of target variables of the $M$ climate models
($T$ denotes the transpose) and $\bm{X} \in \mathbb{R}^{M \times (K + 1)}$ the
design matrix representing the predictors:
\begin{equation}
  \bm{X} =
  \begin{pmatrix}
    1 & x_1^{(1)} & x_1^{(2)} & \cdots & x_1^{(K)} \\
    1 & x_2^{(1)} & x_2^{(2)} & \cdots & x_2^{(K)} \\
    \vdots & \vdots & \vdots & \ddots & \vdots \\
    1 & x_M^{(1)} & x_M^{(2)} & \cdots & x_M^{(K)} \\
  \end{pmatrix}.
  \label{eq:02:design_matrix}
\end{equation}
The entry $x_m^{(k)}$ of this matrix refers to the diagnostic variable of
diagnostic $k$ and climate model $m$. With this notation, the linear
inter-model relationship can be written as
\begin{equation}
  \bm{y} = \bm{X} \bm{b} + \bm{\varepsilon},
  \label{eq:02:multivariate_linear_model}
\end{equation}
where $\bm{b} = \left( b_0, b_1, \ldots, b_K \right)^T \in \mathbb{R}^{(K +
  1)}$ is the vector of linear coefficients (with intercept $b_0$) and
$\bm{\varepsilon} = \left( \varepsilon_1, \varepsilon_2, \ldots, \varepsilon_M
\right)^T \in \mathbb{R}^M$ a vector of independent random variables
representing the noise in the target variable. \Cref{fig:02:mder} shows a
schematic that illustrates this linear relationship (gray surface) for the
different climate models (colored circles) for two diagnostics ($K = 2$). Using
ordinary least squares regression, the estimated linear coefficients
$\hat{\bm{b}}$ are given by
\begin{equation}
  \hat{\bm{b}} = \left( \bm{X}^T \bm{X} \right)^{-1} \bm{X}^T \bm{y},
  \label{eq:02:linear_coefficients}
\end{equation}
where the exponent \enquote{$-1$} denotes the inverse matrix. Since this
definition works for any number of diagnostics $K$, it can also be used to
calculate the linear coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$
(slope) that define emergent relationships (see
\cref{eq:02:linear_regression_y}).

>>>>

A possible way out is given by numerical climate models, which offer the
possibility to simulate the Earth's climate on a computer. To efficiently
replicate the vastly complex Earth system with finite computational resources,
climate models divide the Earth into a set of grid cells. The typical size of
an atmospheric grid cell in modern-day global climate models is about $100
\unit{km}$ along the horizontal dimensions (latitude and longitude) and $1
\unit{km}$ along the vertical dimension (pressure level or height). In
addition, also the temporal evolution of the Earth system is discretized using
time steps that are typically about $30 \unit{min}$ long in modern-day global
climate models. For each grid cell and time step, usually a single value per
model variable is given. Examples for such variables provided by a climate
model are the atmospheric prognostic variables velocity (horizontal and
vertical), temperature, specific humidity, pressure, and density. These
prognostic variables are related to each other via the \emph{primitive
  equations}, a set of partial differential equations that can be derived from
the conservation of momentum, mass, energy, and moisture \autocite{Holton2004}.
To progress further in time, these primitive equations are solved numerically
by the dynamical core of the climate models, which eventually describes the
Earth's large-scale atmospheric motions. Similar to this, many other processes
of the Earth system are simulated by the climate models using other fundamental
physical laws and principles. As opposed to these calculated variables that
form the output of climate models, the corresponding input is mainly given by
the radiative forcing.

@@

Simulations from climate model ensembles of \ac{CMIP} allow us to assess future
climate change in a consistent and transparent way. Especially the
\ac{ScenarioMIP} experiments can give valuable insights into possible
developments of the Earth system by providing future \emph{climate
  projections}. In contrast to climate predictions, climate projections run
over multiple decades and depend upon the future scenario considered, which are
based on assumptions that may or may not turn out to be correct. On the
contrary, climate predictions are attempts to predict the actual evolution of
the climate on much shorter time scales from seasons to years. Similar to any
other scientific experiment, climate model projections suffer from associated
uncertainties. There are three major sources of uncertainties in climate model
projections we can distinguish: natural variability, climate response
uncertainty, and emission uncertainty \autocite{Hawkins2009, Hawkins2010}.
\Cref{fig:02:sources_of_uncertainty} shows these three sources for the
projected global mean surface temperature anomaly over the \nth{21} century.

@@

\emph{Natural variability} is connected to the chaotic nature of the Earth
system that arises from complex interactions between the ocean, atmosphere,
land, biosphere, and cryosphere \autocite{Cubasch2013}. It constitutes a
fundamental limit of how precisely we can project the future climate since it
is inherent in the Earth system and cannot be eliminated by more knowledge and
more advanced climate models. Natural variability is more relevant on regional
and local scales than on continental or global scales. Further contributions to
natural variability on longer time scales come from phenomena like the
\ac{ENSO} or the \ac{NAO} and from other events like volcanic eruptions and
variations in the solar activity. Natural variability can be seen as the
\emph{noise} in the climate record as opposed to the anthropogenic
\emph{signal} \autocite{Cubasch2013}.

@@

A commonly used distance metric to measure model performance $D_m$ and model
interdependence $S_{mn}$ is the \ac{RMSE}, but others are possible
\autocite{Knutti2017a}. The metrics are evaluated on a set of past or
present-day diagnostics and variables, whose choice is crucial for the
weighting scheme. A helpful strategy for this is to focus on addressing the
question \enquote{which climate model is adequate for predicting the target
  variable $y$?} instead of trying to answer the question \enquote{which
  climate model is the best?} \autocite{Parker2009}. Thus, diagnostics and
variables are chosen that are relevant for the projection of the target
variable \autocite{Knutti2017a}. In practice, this choice is either based on
expert judgment about relevant processes, on emergent relationships (see
\cref{subsec:02:emergent_constraints}), or on multivariate regression models
\commentcite{Karpechko2013, Sanderson2015a}{see \cref{subsec:02:mder}}. It
might also be beneficial to use different diagnostics for the calculations of
the performance and interdependence metrics \autocite{Merrifield2020} and/or to
remove selected diagnostics based on their mutual correlation
\autocite{Lorenz2018}.

The constants $\sigma_D$ and $\sigma_S$ determine how strongly the climate
models' performance and interdependence are weighted \autocite{Knutti2017a}.
Small values of the performance parameter $\sigma_D$ lead to an aggressive
weighting with only a few climate models receiving a majority of the weight,
while large values of $\sigma_D$ result in an equal weighting. For the
interdependence parameter $\sigma_S$, this is slightly different: Here, small
(all climate models are independent) and large (all climate models are
dependent) values lead to an almost equal weighting. Thus, an optimal choice
for $\sigma_D$ and $\sigma_S$ is crucial. A useful tool to estimate these
optimal parameters is the \emph{leave-one-model-out \ac{CV}} approach, which is
also known as \emph{pseudo-reality}, \emph{model-as-truth}, or \emph{perfect
  model} setup \autocite{Elia2002, Karpechko2013}. For this, a single climate
model is removed from the multi-model ensemble and treated as observation
(\emph{pseudo-observation}). Then, a weighted \ac{MMM} with weights computed
from the updated model ensemble is calculated, which gives a prediction for the
\enquote{true} climate model. This allows a simple quantitative assessment of
the weighting scheme by calculating the \ac{RMSE} between the prediction and
the known ground truth of the pseudo-observation, which is also known as
\emph{\ac{RMSEP}}. The whole process is repeated for every climate model of the
ensemble to get a statistical distribution of \acp{RMSEP}. Finally, different
\ac{RMSEP} distributions calculated from different parameters $\sigma_D$ and
$\sigma_S$ can be assessed using specific criteria to find optimal values for
$\sigma_D$ and $\sigma_S$ \autocite{Knutti2017a}. Furthermore, the
leave-one-model-out \ac{CV} approach can be used to evaluate different climate
model weightings schemes (including the unweighted \ac{MMM}) and compare them
against each other.

@@

The concept of \ac{MDER} is mathematically similar to the concept of emergent
constraints. In a first step, an inter-model relationship between the target
variable and the process-relevant diagnostics is used to fit a multivariate
linear regression model. Let $\bm{y} = \left( y_1, y_2, \ldots, y_M \right)^T
\in \mathbb{R}^M$ be the vector of target variables of the $M$ climate models
($T$ denotes the transpose) and $\bm{X} \in \mathbb{R}^{M \times (K + 1)}$ the
design matrix representing the predictors:
\begin{equation}
  \bm{X} =
  \begin{pmatrix}
    1 & x_1^{(1)} & x_1^{(2)} & \cdots & x_1^{(K)} \\
    1 & x_2^{(1)} & x_2^{(2)} & \cdots & x_2^{(K)} \\
    \vdots & \vdots & \vdots & \ddots & \vdots \\
    1 & x_M^{(1)} & x_M^{(2)} & \cdots & x_M^{(K)} \\
  \end{pmatrix}.
  \label{eq:02:design_matrix}
\end{equation}
The entry $x_m^{(k)}$ of this matrix refers to the diagnostic variable of
diagnostic $k$ and climate model $m$. With this notation, the linear
inter-model relationship can be written as
\begin{equation}
  \bm{y} = \bm{X} \bm{b} + \bm{\varepsilon},
  \label{eq:02:multivariate_linear_model}
\end{equation}
where $\bm{b} = \left( b_0, b_1, \ldots, b_K \right)^T \in \mathbb{R}^{(K +
  1)}$ is the vector of linear coefficients (with intercept $b_0$) and
$\bm{\varepsilon} = \left( \varepsilon_1, \varepsilon_2, \ldots, \varepsilon_M
\right)^T \in \mathbb{R}^M$ a vector of independent random variables
representing the noise in the target variable. \Cref{fig:02:mder} shows a
schematic that illustrates this linear relationship (gray surface) for the
different climate models (colored circles) for two diagnostics ($K = 2$). Using
ordinary least squares regression, the estimated linear coefficients
$\hat{\bm{b}}$ are given by
\begin{equation}
  \hat{\bm{b}} = \left( \bm{X}^T \bm{X} \right)^{-1} \bm{X}^T \bm{y},
  \label{eq:02:linear_coefficients}
\end{equation}
where the exponent $-1$ denotes the inverse matrix. Since this definition works
for any number of diagnostics $K$, it can also be used to calculate the linear
coefficients $\hat{b}_0$ (intercept) and $\hat{b}_1$ (slope) that define
emergent relationships (see \cref{eq:02:linear_regression_y}).

### 2021-03-31 ch06_paper_gpp/paper_gpp.tex
Standardized the usage of 'emph' and 'enquote'

Our \ac{GBRT} approach (step 2) is mathematically similar to \ac{MDER}
\autocite{Karpechko2013, Senftleben2020, Wenzel2016a}: We establish a
relationship between process-oriented, physically-relevant diagnostics and
future projections and then utilize this to project today's observed conditions
into the future. We emphasize that the exact nature of this relationship is
strongly dependent on the climate model ensemble considered. Although derived
empirically, we argue that this relationship can still be used to extract
further information from the climate model ensemble for two reasons: First, we
consider process-oriented variables which are physically linked to \ac{GPP},
and second, we ensure its statistical robustness. The latter is achieved by
considering gridded climatological data instead of global or regional means.
The number of points establishing the relation is dramatically increased from
about 50 points for classical emergent constraints to 237852 points (step 2a)
and 16503 (step 2b) in our approach. We validate the relationship by extensive
out-of-sample testing using a leave-one-model-out \ac{CV} approach and randomly
excluding parts of the data prior to training. Furthermore, the large number of
points in combination with the non-linear \ac{GBRT} model allows us to exploit
more information than classical emergent constraints and include non-linear
multivariate relations in the statistical model, including the information
encoded in spatial variation. The prediction phase of our approach (feeding
observation-based data into the trained \ac{GBRT} model) can be interpreted as
an implicit performance weighting: The \ac{GBRT} model creates predictions
based on today's conditions (in form of variables which are physically
connected to \ac{GPP}). However, unlike other performance weighting techniques
\autocite{Knutti2017a, Sanderson2017}, we do not assign constant weights to the
different climate models. On the contrary, since the algorithm works on grid
cell level, it is able to adapt to regional climate characteristics and thus
applies an implicit \emph{local} and \emph{climate-dependent} weighting.
Moreover, the algorithm adds another level of weighting by assigning a relative
relevance score to the different predictors. We emphasize that this is only an
implicit weighting, since it is not possible to extract specific values for the
individual weights due to the complex structure of the \ac{GBRT} model.

>>>>

Our \ac{GBRT} approach (step 2) is mathematically similar to \ac{MDER}
\autocite{Karpechko2013, Senftleben2020, Wenzel2016a}: We establish a
relationship between process-oriented, physically-relevant diagnostics and
future projections and then utilize this to project today's observed conditions
into the future. We emphasize that the exact nature of this relationship is
strongly dependent on the climate model ensemble considered. Although derived
empirically, we argue that this relationship can still be used to extract
further information from the climate model ensemble for two reasons: First, we
consider process-oriented variables which are physically linked to \ac{GPP},
and second, we ensure its statistical robustness. The latter is achieved by
considering gridded climatological data instead of global or regional means.
The number of points establishing the relation is dramatically increased from
about 50 points for classical emergent constraints to 237852 points (step 2a)
and 16503 (step 2b) in our approach. We validate the relationship by extensive
out-of-sample testing using a leave-one-model-out \ac{CV} approach and randomly
excluding parts of the data prior to training. Furthermore, the large number of
points in combination with the non-linear \ac{GBRT} model allows us to exploit
more information than classical emergent constraints and include non-linear
multivariate relations in the statistical model, including the information
encoded in spatial variation. The prediction phase of our approach (feeding
observation-based data into the trained \ac{GBRT} model) can be interpreted as
an implicit performance weighting: The \ac{GBRT} model creates predictions
based on today's conditions (in form of variables which are physically
connected to \ac{GPP}). However, unlike other performance weighting techniques
\autocite{Knutti2017a, Sanderson2017}, we do not assign constant weights to the
different climate models. On the contrary, since the algorithm works on grid
cell level, it is able to adapt to regional climate characteristics and thus
applies an implicit local and climate-dependent weighting. Moreover, the
algorithm adds another level of weighting by assigning a relative relevance
score to the different predictors. We emphasize that this is only an implicit
weighting, since it is not possible to extract specific values for the
individual weights due to the complex structure of the \ac{GBRT} model.
